# Data transformation

The process of importing the data into R was extremely smooth. We downloaded the data from Redfin, extracted it and then read it directly into an R data frame. Afterwards, we changed a few variables to factor and date classes. We then used the dplyr select function to choose only the columns and time range (2018-01-01 to 2021-10-31) we wanted to focus on. Here are the first 6 rows of the dataset.

```{r}
library(tidyverse)
```

```{r}
df <- read.table("USHouse_States.tsv", sep = "\t", header = TRUE)

house <- df %>% select("period_begin", "period_end", "table_id", "state", "state_code", "property_type", "median_sale_price", "median_list_price", "median_ppsf", "median_list_ppsf", "homes_sold", "new_listings", "inventory", "months_of_supply", "avg_sale_to_list", "sold_above_list", "off_market_in_two_weeks", "parent_metro_region")%>%
  mutate(period_begin=lubridate::ymd(period_begin))%>%
  mutate(period_end=lubridate::ymd(period_end))%>%
  filter(property_type=="All Residential",
         period_begin>=as.Date("2018-01-01"),
         period_end>=as.Date("2018-01-31"))%>%
  arrange(desc(period_begin))

```

```{r}
head(house)
```



